Developing Observational Health Research Capacity in the Julia Ecosystem

### 2023-12-10 paper.tex
Working on making major revisions

\usepackage{listings}

@@

As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.

@@

% DONE
JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.

@@

% DONE:

@@

\begin{lstlisting}[language=Julia,numbers=none]
import DBConnector: 
  DBConnection
import HealthSampleData: 
  Eunomia
import OMOPCDMCohortCreator as OCC

@@

occ.GenerateTables(conn)
\end{lstlisting}

@@

for query in split(sql, ";")[1:end-1]
  DBI.execute(conn, sub_query)
end
\end{minted}

@@

\begin{lstlisting}[language=Julia,numbers=none]
C = GetCohortSubjects(1, conn)
rename!(C, :subject_id => :person_id)

@@

C_age_group = occ.GetPatientAgeGroup(
  C.person_id, 
  conn; 
)
\end{lstlisting}

@@

\begin{lstlisting}[language=Julia,numbers=none]
import DataFrames as DF

@@

C_groups = DF.combine(C_groups, DF.nrow => :count)
\end{lstlisting}

@@

\begin{lstlisting}[language=Julia,numbers=none]
import Base:
  Fix2

@@

CGetPatientRace = Fix2(GetPatientRace, c)
CGetPatientGender = Fix2(GetPatientGender, c)
CGetPatientAgeGroup = Fix2(GetPatientAgeGroup, c)
\end{lstlisting}

@@

\begin{lstlisting}[language=Julia,numbers=none]
import Chain:
  @chain

CofactorPatients(patients) = @chain patients begin
    CGetPatientRace_fixed
    CGetPatientGender_fixed
    CGetPatientAgeGroup_fixed
end
\end{lstlisting}

\begin{lstlisting}[language=Julia,numbers=none]
CharacterizePatients(patients) = @chain patients begin
    _[:, DF.Not(:person_id)]
    DF.groupby(_, names(_))
    DF.combine(_, DF.nrow => :count)
    occ.ExecuteAudit
end
\end{lstlisting}

@@

\begin{lstlisting}[language=Julia,numbers=none]
RunStudy(patients) = @chain patients begin
    CofactorPatients
    CharacterizePatients
end
\end{lstlisting}

@@

\begin{lstlisting}[language=Julia,numbers=none]
import Distributed:
  @distributed

@@

vcat(result...) |> audit_patient_groups
\end{lstlisting}

@@

\begin{lstlisting}[language=Julia,numbers=none]
  GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
\end{lstlisting}

\begin{lstlisting}[language=Julia,numbers=none]
  SELECT
    "PERSON_2"."person_id",
    (CASE 
      WHEN ("PERSON_2"."age" < 41) 
        THEN '0 - 40' 
      WHEN ("PERSON_2"."age" < 81) 
        THEN '41 - 80' END) AS "age_group"
  FROM (
    SELECT
      "PERSON_1"."person_id",
      (2023 - "PERSON_1"."year_of_birth") AS "age"
    FROM "PERSON" AS "PERSON_1"
    WHERE ("PERSON_1"."person_id" IN (1))
  ) AS "PERSON_2"
\end{lstlisting}

>>>>

@@

As the OMOP CDM has become a cornerstone in running these widely distributed network studies, special attention to developing tools around handling OMOP CDM data is necessary to meaningfully contribute within the observational health research space.
% TODO: Move this to the section: Status of Conducting Observational Health Research in Julia
The OMOP CDM itself does not require a specific technology to work with the data stored in this standard.
It features a person-centric design where each domain records personal identity while prioritizing data protection through the limiting of information that could endanger patient anonymity.

@@

JuliaHealth is an open-science collective dedicated to bringing the Julia programming language to various avenues of health research domains including, but not limited to, observational health, medical imaging, standards interoperability and more.
It leverages the strengths of the Julia language to introduce new methods into these spaces, augment existing workflows, and provide a community structure for health researchers within the global Julia ecosystem.
Eventually, JuliaHealth will host a variety of subecosystems and associated resources that will enable novel research to be conducted within Julia.

@@

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import DBConnector: 
  DBConnection
import HealthSampleData: 
  Eunomia
import OMOPCDMCohortCreator as OCC

@@

occ.GenerateTables(conn)
\end{minted}

@@

for query in split(sql, ";")[1:end-1]
  DBI.execute(conn, query)
end
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
C = GetCohortSubjects(1, conn)
rename!(C, :subject_id => :person_id)

@@

C_age_group = occ.GetPatientAgeGroup(
  C.person_id, 
  conn; 
)
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import DataFrames as DF

@@

C_groups = DF.combine(C_groups, DF.nrow => :count)
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import Base:
  Fix2

@@

CGetPatientRace = Fix2(GetPatientRace, c)
CGetPatientGender = Fix2(GetPatientGender, c)
CGetPatientAgeGroup = Fix2(GetPatientAgeGroup, c)
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import Chain:
  @chain

CofactorPatients(patients) = @chain patients begin
    CGetPatientRace_fixed
    CGetPatientGender_fixed
    CGetPatientAgeGroup_fixed
end
\end{minted}

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
CharacterizePatients(patients) = @chain patients begin
    _[:, DF.Not(:person_id)]
    DF.groupby(_, names(_))
    DF.combine(_, DF.nrow => :count)
    occ.ExecuteAudit
end
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
RunStudy(patients) = @chain patients begin
    CofactorPatients
    CharacterizePatients
end
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
import Distributed:
  @distributed

@@

vcat(result...) |> audit_patient_groups
\end{minted}

@@

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
  GetPatientAgeGroup([1]; age_groupings = [[0, 40], [41, 80]])
\end{minted}

\begin{minted}[breaklines,escapeinside=||,mathescape=true, numbersep=3pt, gobble=2, frame=lines, fontsize=\scriptsize, framesep=2mm]{julia}
  SELECT
    "PERSON_2"."person_id",
    (CASE 
      WHEN ("PERSON_2"."age" < 41) 
        THEN '0 - 40' 
      WHEN ("PERSON_2"."age" < 81) 
        THEN '41 - 80' END) AS "age_group"
  FROM (
    SELECT
      "PERSON_1"."person_id",
      (2023 - "PERSON_1"."year_of_birth") AS "age"
    FROM "PERSON" AS "PERSON_1"
    WHERE ("PERSON_1"."person_id" IN (1))
  ) AS "PERSON_2"
\end{minted}
